Reproducible Research: Peer Assessment 1
=========================================
####(... or what else can we see about anonymous' walking data gathering)

## Loading and preprocessing the data
Packages used:
```{r message=F, warning=F}
library(lubridate)
library(dplyr)
library(lattice)
library(ggplot2)
```

Dataset loading + 
        day_of_week/month_of_year extracting and assigning to columns 
        'dowk', "mofyear"
```{r}
hishersteps <- read.csv("activity.csv")
hishersteps <- mutate(hishersteps, dowk = wday(hishersteps$date))
hishersteps <- mutate(hishersteps, mofyear = month(ymd(date))) 
```

## What is mean total number of steps taken per day?

- The total number of steps per day (ignoring missing values)
```{r}
sum(hishersteps$steps, na.rm = TRUE)
```

- Histogram of the total number of steps taken each day (ignoring missing values)
```{r}
byday <- group_by(hishersteps, mofyear, dowk)
par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
hist(subset(byday, (mofyear == 10 & steps > 0))$steps, 
                col = "green", xlab = "# of steps in mean_minus_14days chunks",
                main = "Histogram of October steps", 
                breaks = mean(hishersteps$steps, na.rm = TRUE)-14)
hist(subset(byday, (mofyear == 11 & steps > 0))$steps, 
                col = "green", xlab = "# of steps in mean_minus_14days chunks",
                main = "Histogram of November steps", 
                breaks = mean(hishersteps$steps, na.rm = TRUE)-14)
```

- The mean and median of total steps per day (ignoring missing values)
```{r}
meanNmed <- summarize(hishersteps, stepsMean = mean(hishersteps$steps, na.rm = TRUE), 
                                   stepsMedian =median(hishersteps$steps, na.rm = TRUE))
head(meanNmed)
```


## What is the average daily activity pattern?
- time series plot of the 5-minute interval (x-axis) and the average number of
steps taken, averaged across all days (y-axis) 
```{r}
byinterval <- group_by(byday, interval, date)
byinterval2 <- summarize(byinterval, meanOfday = mean(steps, na.rm = TRUE))
xyplot(meanOfday ~ interval, data = byinterval2, na.rm = TRUE, 
                                col = rgb(0, 0, 0, 0.50),
                                main = "Mean of steps by interval by day",
                                xlab = "5 min. Intervals",
                                ylab = "Means of day's steps",
                                breaks = 10)



```

- Which 5-minute interval... contains the maximum number of steps?

```{r}
maxSteps <-max(as.numeric(byinterval2$meanOfday), na.rm = TRUE)
theInterval <- filter(byinterval2, meanOfday == maxSteps)
theInterval$interval
```

## Imputing missing values
- Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(hishersteps$steps))
```
- Devise a strategy for filling in all of the missing values in the dataset.
```{r}
minSteps <-min(as.numeric(byinterval2$meanOfday), na.rm = TRUE)
fillerStep <- (minSteps + maxSteps)/2
```
- Create a new dataset with filled missing values
```{r}
hisherfilled <- hishersteps
hisherfilled[,1:1][is.na(hisherfilled[,1:1])] <-fillerStep
```
- Histogram with NAs replaced
```{r}
byday2 <- group_by(hisherfilled, mofyear, dowk)
#head(byday2)
par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
hist(subset(byday2, (mofyear == 10 & steps > 0))$steps, 
                col = "green", xlab = "# of steps in mean_minus_14days chunks",
                main = "Histogram of October steps (NAs replaced)", 
                breaks = mean(hishersteps$steps, na.rm = TRUE)-14)
hist(subset(byday2, (mofyear == 11 & steps > 0))$steps, 
                col = "green", xlab = "# of steps in mean_minus_14days chunks",
                main = "Histogram of November steps (NAs replaced)", 
                breaks = mean(hishersteps$steps, na.rm = TRUE)-14)
```

- What is the impact of imputing missing data on total daily number of steps?
1. Number of steps with missing data
```{r}
sum(hishersteps$steps, na.rm = TRUE)
```
2. Number of steps after filling missing rows with half maxSteps value
```{r}        
sum(hisherfilled$steps, na.rm = TRUE)
```


## Are there differences in activity patterns between weekdays and weekends?
